Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-06-16 manuscript.tex
group_split_sensor_reduce

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{body position, motor development, everyday experiences, sitting, machine learning\newline\indent Word count: 11,995}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}
\raggedbottom
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Long-form recording of infant body position in the home using wearable inertial sensors},
  pdfauthor={John M. Franchak1, Maximilian Tang1, Hailey Rousey1, \& Chuan Luo1},
  pdflang={en-EN},
  pdfkeywords={body position, motor development, everyday experiences, sitting, machine learning},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

From moment to moment, infants' movements facilitate and constrain how they can interact with their surroundings. Changes in \emph{body position}---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a caregiver---have immediate consequences for vision, object exploration, and social interaction. When sitting and upright, infants have a better view of faces and distant objects compared to their view while in a prone position (Franchak et al., 2018; Kretch et al., 2014; Luo \& Franchak, 2020). Learning to sit facilitates object exploration (Soska \& Adolph, 2014). Upright walking changes how infants and caregivers interact; walking infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position (Chen et al., 2022; Karasik et al., 2011). As infants grow older and acquire new abilities, such as independent sitting and walking, they spend more time sitting and upright and less time held, supine, and prone (Adolph \& Tamis-LeMonda, 2014; Franchak et al., 2018; Franchak, 2019; Thurman \& Corbetta, 2017). Caregivers provide different learning input to infants depending on infants' body position (Karasik et al., 2014; Kretch et al., 2022; West \& Iverson, 2021). For example, infants' walking movements spur caregivers to talk about actions rather than just describe objects. Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning (Franchak, 2020).

In this paper, we present an inertial sensing method to classify infant body position from moment to moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders. We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions (Airaksinen et al., 2022, 2020; Franchak et al., 2021), the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}.

@@

The LENA® recorder is a commercial device worn in a custom shirt pocket; the recorder has sufficient battery life and storage to record for an entire day. Closed-source LENA® algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the infant. Other long-form audio methods rely on custom-built recorders (Wass et al., 2022), apply alternative classification algorithms to LENA® data (Micheletti et al., 2022; Räsänen et al., 2020), or manually transcribe audio recorded by LENA® devices to improve accuracy or identify behaviors beyond the built-in categories (Bergelson, Casillas, et al., 2019; Mendoza \& Fausey, 2021).

Long-form audio recordings have had a transformational impact on language development research by allowing researchers to characterize opportunities for learning in daily life. For example, measuring the amount of speech heard by infants in the home (Weisleder \& Fernald, 2013) or in a daycare setting (Perry et al., 2018) revealed how individual differences in speech input predict later vocabulary. Full-day language recording synchronized with other data sources allows researchers to identify how auditory input and vocal production interact with other processes. Beyond individual differences in aggregated data, long-form recordings reveal the temporal schedule of experiences. For example, infants' daily experiences hearing music are clustered in time, with ``bursty'' episodes of music presence separated by relatively long periods during which music is absent (Mendoza \& Fausey, 2022). In addition, synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences. Linking LENA® speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during periods that they were restrained in devices such as swings, exersaucers, and car seats (Malachowski et al., 2023). A custom-built wearable ECG and audio recorder allowed Wass et al. (2022) to discover that infant arousal increases the likelihood of infant vocalization across the day.

We identified five key features of long-form audio methods that should be replicated in analogous studies of motor behavior. First, wearable audio recorders are \emph{mobile}. Measurement is not limited to a particular room because the recording device travels with the participant. Data are recorded to onboard device memory, so participants do not need to be in range of a receiver. Second, wearable audio recording is more \emph{unobtrusive}. Participants' reactivity to observation, such as from a video camera, may influence behavior more compared with a sensor that records only motion or audio data. For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA® device (Bergelson, Amatuni, et al., 2019). Third and fourth, recordings capture \emph{real-time data} over a \emph{full day}. The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants. Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual (Malachowski et al., 2023; e.g., Wass et al., 2022). Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) (Kadooka et al., 2021, April; Tamis-LeMonda et al., 2018). ``Burstiness'' of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent (Barbaro \& Fausey, 2022; Warlaumont et al., 2021). Fifth, \emph{automatic classification} means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.

But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased. An independent assessment of the LENA® algorithms found mixed results about the accuracy of different measurements. For example, correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively (Cristia et al., 2020). For other metrics, such as the number of ``conversational turns'' between the child and communicative partners, agreement was poor (\emph{r} = .364). Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day. In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior.

@@

Video observation is the most common way of measuring infant motor behavior in home recordings. Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session (Chen et al., 2022; Herzberg et al., 2021; Karasik et al., 2011). The primary advantage of video recording is that it captures real-time behavior. Infants' body position, locomotion, and reaching comprise events that occur on the timescale of seconds, so standard video recording is adequate to score gross motor behavior. However, requiring an experimenter to operate a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place. Moreover, video observation cannot easily scale to long durations or large numbers of participants. Logistically, an experimenter cannot follow behind infants to record their behavior from morning to night (and were they to do so, they would likely alter infants' and caregivers' behavior). Typical video recording sessions last 45-120 minutes (Chen et al., 2022; Herzberg et al., 2021; Karasik et al., 2011), far short of capturing the variety of activities across the full daily routine. Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants. Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old (Galland et al., 2012) could take 22-55 hours of labor to fully annotate.

In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation. Diary studies provide caregivers with logs or structured interviews to estimate from memory how much time infants spend in particular activities (Karasik et al., 2022; Majnemer \& Barr, 2005). Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day (Franchak, 2019; Kadooka et al., 2021, April). Although the EMA responses are valuable in aggregate, survey methods lack the real-time temporal resolution to describe moment-to-moment changes in behavior. For diaries and interviews, limits on caregivers' memory mean that they will report what was most frequent, but cannot remember events that happen on the scale of seconds and minutes. At best, EMA surveys prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent. Thus, despite being a useful tool for estimating broad developmental changes and individual differences in infants' motor experiences, survey methods are not suited for capturing within-participant temporal dynamics.

@@

Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home (Barbaro, 2019; Bruijns et al., 2020; Cliff et al., 2009; Lobo et al., 2019). Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior. Many commercially-available and inexpensive IMUs have \textgreater{} 12 hour battery life with onboard storage to record \emph{real-time}, \emph{full-day} motion data at a sampling rate that is higher than typical video (e.g., 50-100 Hz). Indeed, past work has successfully recorded the rate of leg kicks (e.g., Deng et al., 2019) and activity intensity (Schneller et al., 2017) in infants and children across multiple days.

The open question is whether \emph{automatic classification} is sufficiently accurate to measure movement categories that are relevant to developmental and clinical research, and whether measurements continue to be accurate over long recording periods. IMUs typically contain accelerometers that measure linear acceleration paired with gyroscopes that measure angular acceleration. Unlike motion tracking systems that might be used in a lab, IMUs do not provide data about the position of the body in space. Thus, data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration time series) into meaningful categories (e.g., supine, prone, sitting, upright). The difficulty of the classification task depends on the categories of interest. More basic aspects of movement, such as overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity (Armstrong et al., 2019; e.g., Hager et al., 2017).

Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is too complex for cut-point definitions to be accurate. First, the magnitude of movement can vary greatly \emph{within} a body position. An upright infant can be standing still or can be walking briskly across the room. A prone infant can be stationary in ``tummy time'', or they can crawl in a myriad of ways (Adolph et al., 1998). Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts. Infants can sit on the floor in a tripod position with support from an arm, in a ``V'' position with legs fully extended, or in a ``W'' position with knees bent. Sitting on a caregiver's lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions. Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso. Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life. Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity (Kwon et al., 2019; Patel et al., 2019). Although Greenspan et al. (2021) were able to use cut-point definitions about infant trunk angle to categorize body positions, this method was not able to differentiate times that infants were held by caregivers off the ground. Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults (Arif \& Kattan, 2015; Preece et al., 2009), children (Nam \& Park, 2013; Ren et al., 2016; Stewart et al., 2018), and infants (Airaksinen et al., 2020; Franchak et al., 2021; Yao et al., 2019). Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest model, that can be later used to predict categories from IMU data that was not annotated. Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotations.

Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data. Airaksinen et al. (2020) tested 4- to 8-month-olds in a laboratory visit, and found 95\% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers). Using a wider age range of 6-18 months, Franchak et al. (2021) found 98\% accuracy (\emph{kappa} = 95\%) in a laboratory validation study in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding. Most recently, Airaksinen et al. (2022) conducted a validation study of body position classification in either a home or clinic testing in 4- to 19-month-olds, refining their previous method to detect moments that infants were carried by caregivers. Classification accuracy did not vary between lab and home settings, and was generally high (95\%, \emph{kappa} = .93). Although all three studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.

@@

Accordingly, the overarching goal of the current study is to test the validity of long-form body position recording in the home during unsupervised, everyday behavior. Supervised recordings from past work (Airaksinen et al., 2022, 2020; Franchak et al., 2021), whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors. Prior work focused on ``free play'', in which caregivers were asked to play with the infant without restraining the infant or otherwise shaping how they could move. However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position. For example different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to scored as the same behavior. Moreover, classifiers must be able to detect new variants of behaviors that might arise over the course of a real day; it is impossible for researchers to gather training data for every possible variation that might occur. Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time? Assessing the validity of temporally \emph{distal} periods---when infants and caregivers are unsupervised and free to follow their everyday routines---is a crucial step to establish whether automatic classification can be used to measure body position across a day.

In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 34 testing sessions from 22 infants aged 4-14 months. Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position. A \emph{proximal comparison} period began when participants received the equipment and completed a guided phone call during which caregivers were asked to elicit different body positions based on prompts from the experimenter. This ``semi-supervised'' period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play while they received instructions from the experimenter. The \textbf{first goal} of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes. Past work found better performance using ``individual models''---models that were trained on one participant's data to predict their later behavior---compared with a ``group model'' that aggregated data from all infants to create a single body position classifier (Franchak et al., 2021), so we compared both modeling approaches in the current investigation.

@@

After the distal comparison period when video recording ceased, we asked caregivers to have infants wear the IMUs for the rest of the day until their regular bedtime, creating the \textbf{first real-time, full-day dataset of infant body position}. Interpreting such data required caregivers to log when infants napped, when they removed the sensor garment for diaper changes or other reasons, and when infants went to bed at the end of the day. Thus, the \textbf{third goal} of the study was to examine the quality of the full-day data. Could body position be successfully recorded over the desired period? If full-day classifications of infant behavior are accurate, they should demonstrate convergent validity with other methods. Thus, we determined whether full-day body position measurements conformed to expected age differences in body position that have been identified from previous video and survey methods. Based on past results (Franchak, 2019), infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).

@@

Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study. Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of 34 sessions, with 14 sessions from younger infants and 20 sessions from older infants. Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.2) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7). The total number of recording sessions (34) exceeded the number of sessions employed in comparable past work: 10 in @Nam and Park (2013), 15 in Franchak et al. (2021), 22 in Airaksinen et al. (2020), and 33 in Yao et al. (2019).

>>>>

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{body position, motor development, everyday experiences, sitting, machine learning\newline\indent Word count: 11,995}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}
\raggedbottom
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[style=apa,]{biblatex}
\addbibresource{master.bib}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Long-form recording of infant body position in the home using wearable inertial sensors},
  pdfauthor={John M. Franchak1, Maximilian Tang1, Hailey Rousey1, \& Chuan Luo1},
  pdflang={en-EN},
  pdfkeywords={body position, motor development, everyday experiences, sitting, machine learning},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

From moment to moment, infants' movements facilitate and constrain how they can interact with their surroundings. Changes in \emph{body position}---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a caregiver---have immediate consequences for vision, object exploration, and social interaction. When sitting and upright, infants have a better view of faces and distant objects compared to their view while in a prone position \autocite{Freeplay,LuoFranchak2020,CWW}. Learning to sit facilitates object exploration \autocite{SoskaAdolph2014}. Upright walking changes how infants and caregivers interact; walking infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position \autocite{ChenSchneider2022,Karasik2011}. As infants grow older and acquire new abilities, such as independent sitting and walking, they spend more time sitting and upright and less time held, supine, and prone \autocite{Survey,Thurman2017,Freeplay,AdolphCTL2014}. Caregivers provide different learning input to infants depending on infants' body position \autocite{KretchKoziol2022,WestIverson2021,Karasik2014}. For example, infants' walking movements spur caregivers to talk about actions rather than just describe objects. Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning \autocite{CurrentOpinion}.

In this paper, we present an inertial sensing method to classify infant body position from moment to moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders. We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions \autocite{FranchakScott2021,AiraksinenRasanen2020,AiraksinenGallen2022}, the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}.

@@

The LENA® recorder is a commercial device worn in a custom shirt pocket; the recorder has sufficient battery life and storage to record for an entire day. Closed-source LENA® algorithms analyze the audio recordings to provide automatic counts of useful metrics, such as the number of words spoken by adults in the vicinity of the infant. Other long-form audio methods rely on custom-built recorders \autocite{WassPhillips2022}, apply alternative classification algorithms to LENA® data \autocite{RasanenSeshadri2020,MichelettiYao2022}, or manually transcribe audio recorded by LENA® devices to improve accuracy or identify behaviors beyond the built-in categories \autocite{Bergelson2019Corpus,MendozaFausey2021}.

Long-form audio recordings have had a transformational impact on language development research by allowing researchers to characterize opportunities for learning in daily life. For example, measuring the amount of speech heard by infants in the home \autocite{Weisleder2013} or in a daycare setting \autocite{PerryPrince2018} revealed how individual differences in speech input predict later vocabulary. Full-day language recording synchronized with other data sources allows researchers to identify how auditory input and vocal production interact with other processes. Beyond individual differences in aggregated data, long-form recordings reveal the temporal schedule of experiences. For example, infants' daily experiences hearing music are clustered in time, with ``bursty'' episodes of music presence separated by relatively long periods during which music is absent \autocite{MendozaFausey2022}. In addition, synchronizing audio recordings with other data sources extends researchers' ability to characterize daily experiences. Linking LENA® speech measurements to repeated, time-stamped text-message surveys about infant device placement revealed that infants heard less caregiver speech during periods that they were restrained in devices such as swings, exersaucers, and car seats \autocite{MalachowskiSalo2023}. A custom-built wearable ECG and audio recorder allowed \textcite{WassPhillips2022} to discover that infant arousal increases the likelihood of infant vocalization across the day.

We identified five key features of long-form audio methods that should be replicated in analogous studies of motor behavior. First, wearable audio recorders are \emph{mobile}. Measurement is not limited to a particular room because the recording device travels with the participant. Data are recorded to onboard device memory, so participants do not need to be in range of a receiver. Second, wearable audio recording is more \emph{unobtrusive}. Participants' reactivity to observation, such as from a video camera, may influence behavior more compared with a sensor that records only motion or audio data. For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA® device \autocite{Bergelson2019Input}. Third and fourth, recordings capture \emph{real-time data} over a \emph{full day}. The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants. Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual \autocites[e.g.,][]{WassPhillips2022,MalachowskiSalo2023}. Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) \autocite{KadookaCaufield2021,Tamis-LeMondaCustode2018}. ``Burstiness'' of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent \autocite{BarbaroFausey2022,WarlaumontSobowale2021}. Fifth, \emph{automatic classification} means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.

But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased. An independent assessment of the LENA® algorithms found mixed results about the accuracy of different measurements. For example, correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively \autocite{CristiaLavechin2020}. For other metrics, such as the number of ``conversational turns'' between the child and communicative partners, agreement was poor (\emph{r} = .364). Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day. In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior.

@@

Video observation is the most common way of measuring infant motor behavior in home recordings. Most often, an experimenter with a handheld camera follows infants from room to room to ensure that their movements are visible throughout the recording session \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021}. The primary advantage of video recording is that it captures real-time behavior. Infants' body position, locomotion, and reaching comprise events that occur on the timescale of seconds, so standard video recording is adequate to score gross motor behavior. However, requiring an experimenter to operate a camera is obtrusive, whereas relying on a stationary camera means that infants will be absent from view as they move from place to place. Moreover, video observation cannot easily scale to long durations or large numbers of participants. Logistically, an experimenter cannot follow behind infants to record their behavior from morning to night (and were they to do so, they would likely alter infants' and caregivers' behavior). Typical video recording sessions last 45-120 minutes \autocite{Karasik2011,ChenSchneider2022,HerzbergFletcher2021}, far short of capturing the variety of activities across the full daily routine. Even if full day videos were available, the lack of suitable automatic classification tools means that the human cost of annotation makes it difficult for video methods to scale to testing large numbers of participants. Our annotation of body position takes approximately 2-5 hours to complete for every hour of video (depending on how often infants switch positions), meaning that a full ``waking day'' of approximately 11 hours for a 12-month-old \autocite{GallandTaylor2012} could take 22-55 hours of labor to fully annotate.

In contrast, survey methods such as daily diaries/inventories or ecological momentary assessment (EMA) are mobile, unobtrusive, can be applied across an entire day, and do not need laborious annotation. Diary studies provide caregivers with logs or structured interviews to estimate from memory how much time infants spend in particular activities \autocite{MajnemerBarr2005,KarasikKuchirko2022}. Ecological momentary assessment uses text-message or app-based notifications to prompt caregivers to make repeated estimates of infants' behavior over the course of a day \autocite{Survey,KadookaCaufield2021}. Although the EMA responses are valuable in aggregate, survey methods lack the real-time temporal resolution to describe moment-to-moment changes in behavior. For diaries and interviews, limits on caregivers' memory mean that they will report what was most frequent, but cannot remember events that happen on the scale of seconds and minutes. At best, EMA surveys prompt caregivers to make hourly observations; increasing the number of surveys per day would be too burdensome for the respondent. Thus, despite being a useful tool for estimating broad developmental changes and individual differences in infants' motor experiences, survey methods are not suited for capturing within-participant temporal dynamics.

@@

Measuring infant movement with inertial movement units (IMUs) is a promising avenue for long form recordings of motor behavior in the home \autocite{CliffReilly2009,Barbaro2019,BruijnsTruelove2020,LoboHall2019}. Lightweight sensors (10-30 g) can be embedded in garments to make recordings fully \emph{mobile}, and they are \emph{unobtrusive} because they do not require a researcher's presence nor do they record sensitive audio/video that might influence participants' behavior. Many commercially-available and inexpensive IMUs have \textgreater{} 12 hour battery life with onboard storage to record \emph{real-time}, \emph{full-day} motion data at a sampling rate that is higher than typical video (e.g., 50-100 Hz). Indeed, past work has successfully recorded the rate of leg kicks \autocite[e.g.,][]{DengTrujillo-Priego2019} and activity intensity \autocite{SchnellerBentsen2017} in infants and children across multiple days.

The open question is whether \emph{automatic classification} is sufficiently accurate to measure movement categories that are relevant to developmental and clinical research, and whether measurements continue to be accurate over long recording periods. IMUs typically contain accelerometers that measure linear acceleration paired with gyroscopes that measure angular acceleration. Unlike motion tracking systems that might be used in a lab, IMUs do not provide data about the position of the body in space. Thus, data processing algorithms are needed to classify the raw sensor data (i.e., linear and angular acceleration time series) into meaningful categories (e.g., supine, prone, sitting, upright). The difficulty of the classification task depends on the categories of interest. More basic aspects of movement, such as overall activity intensity, can be identified by taking the magnitude of acceleration (irrespective of direction) and applying thresholds or cut-points to define when a movement has occurred at a particular intensity \autocites[e.g.,][]{HagerTilton2017,ArmstrongCovington2019}.

Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is too complex for cut-point definitions to be accurate. First, the magnitude of movement can vary greatly \emph{within} a body position. An upright infant can be standing still or can be walking briskly across the room. A prone infant can be stationary in ``tummy time'', or they can crawl in a myriad of ways \autocite{AdolphVereijken1998}. Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts. Infants can sit on the floor in a tripod position with support from an arm, in a ``V'' position with legs fully extended, or in a ``W'' position with knees bent. Sitting on a caregiver's lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions. Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso. Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life. Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity \autocite{KwonZavos2019,PatelShi2019}. Although \textcite{GreenspanCunha2021} were able to use cut-point definitions about infant trunk angle to categorize body positions, this method was not able to differentiate times that infants were held by caregivers off the ground. Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults \autocite{Preece2009,Arif2015}, children \autocite{NamPark2013,RenDing2016,StewartNarayanan2018}, and infants \autocite{YaoPlotz2019,FranchakScott2021,AiraksinenRasanen2020}. Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest model, that can be later used to predict categories from IMU data that was not annotated. Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotations.

Three prior investigations have used different machine learning techniques to categorize infant body position from IMUs towards the goal of collecting full-day data. \textcite{AiraksinenRasanen2020} tested 4- to 8-month-olds in a laboratory visit, and found 95\% accuracy in distinguishing between body position categories that crawling infants could perform on the floor (excluding times that infants were held by caregivers). Using a wider age range of 6-18 months, \textcite{FranchakScott2021} found 98\% accuracy (\emph{kappa} = 95\%) in a laboratory validation study in categorizing body position that included infants who could both crawl and walk and also included a category for caregiver holding. Most recently, \textcite{AiraksinenGallen2022} conducted a validation study of body position classification in either a home or clinic testing in 4- to 19-month-olds, refining their previous method to detect moments that infants were carried by caregivers. Classification accuracy did not vary between lab and home settings, and was generally high (95\%, \emph{kappa} = .93). Although all three studies yielded promising classification accuracy, accuracy was assessed in brief (15-60 minute) sessions supervised by a researcher, leaving the open question of how well body position classification will scale to testing across an entire day of natural home life.

@@

Accordingly, the overarching goal of the current study is to test the validity of long-form body position recording in the home during unsupervised, everyday behavior. Supervised recordings from past work \autocite{FranchakScott2021,AiraksinenRasanen2020,AiraksinenGallen2022}, whether in the home or in the lab, let researchers set up the situation to encourage or restrict certain behaviors. Prior work focused on ``free play'', in which caregivers were asked to play with the infant without restraining the infant or otherwise shaping how they could move. However, in a real day, non-play activities (e.g., eating lunch in a high chair) create challenging situations for applying automated classification of body position. For example different types of sitting---independently on the floor, supported on a caregiver's lap, or restrained in a high chair---all need to scored as the same behavior. Moreover, classifiers must be able to detect new variants of behaviors that might arise over the course of a real day; it is impossible for researchers to gather training data for every possible variation that might occur. Thus, our central question is whether models trained on video-recorded observations at the beginning of the day generalize to predict behavior at a later time? Assessing the validity of temporally \emph{distal} periods---when infants and caregivers are unsupervised and free to follow their everyday routines---is a crucial step to establish whether automatic classification can be used to measure body position across a day.

In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 34 testing sessions from 22 infants aged 4-14 months. Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position. A \emph{proximal comparison} period began when participants received the equipment and completed a guided phone call during which caregivers were asked to elicit different body positions based on prompts from the experimenter. This ``semi-supervised'' period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play while they received instructions from the experimenter. The \textbf{first goal} of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes. Past work found better performance using ``individual models''---models that were trained on one participant's data to predict their later behavior---compared with a ``group model'' that aggregated data from all infants to create a single body position classifier \autocite{FranchakScott2021}, so we compared both modeling approaches in the current investigation.

@@

After the distal comparison period when video recording ceased, we asked caregivers to have infants wear the IMUs for the rest of the day until their regular bedtime, creating the \textbf{first real-time, full-day dataset of infant body position}. Interpreting such data required caregivers to log when infants napped, when they removed the sensor garment for diaper changes or other reasons, and when infants went to bed at the end of the day. Thus, the \textbf{third goal} of the study was to examine the quality of the full-day data. Could body position be successfully recorded over the desired period? If full-day classifications of infant behavior are accurate, they should demonstrate convergent validity with other methods. Thus, we determined whether full-day body position measurements conformed to expected age differences in body position that have been identified from previous video and survey methods. Based on past results \autocite{Survey}, infants should spend increasingly more time sitting and upright but less time supine over the age range tested (4 to 14 months).

@@

Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study. Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of 34 sessions, with 14 sessions from younger infants and 20 sessions from older infants. Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.2) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7). The total number of recording sessions (34) exceeded the number of sessions employed in comparable past work: 10 in \textcite{NamPark2013}, 15 in \textcite{FranchakScott2021}, 22 in \textcite{AiraksinenRasanen2020}, and 33 in \textcite{YaoPlotz2019}.
